Do syntactic and semantic similarity lead to interference effects?\\ Evidence from self-paced reading and event-related potentials using German

### 2024-11-15 paper/revision2/ResponseToReviewers.tex

\myexternaldocument{elsarticle-template-harv}

\begin{document}
\hfill \today
%\section{}
%\subsection{}

\noindent \textbf{Prof.\ Jarrold}\\
\noindent \textbf{Associate Editor}\\
\noindent \textbf{Journal of Memory and Language}
\vskip 1em

\noindent \textbf{Response letter for Submission JML-24-17, 2nd round of revisions}
\vskip 2em
\noindent Dear Prof. Jarrold, 
\vskip 1em
with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit our second revision of our article titled ``Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German''.

We are grateful for thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
To address the concern that we overstate the implications of our findings, we have clarified at various places in the abstract and discussion of our results that our findings only apply to the present design. Although we have added clarifying text in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), the manuscript length could be shortened by four pages, without compromising on content. We have removed repetitions and have revised unnecessary verbose text passages (e.g., in the description of the SPR results).
We have made our OSF repository public to ensure general accessibility and have updated the link in the manuscript accordingly.

@@

\textcolor{red}{do this actually}

\subsection*{Comment E.2}
\begin{quote}
``The other key points I would like to ask you to pay particular attention to in any response to the current set of reviews are:

\textbf{Reviewer 2 and 3's clear view that you still overstate the implications of your findings in places}.''
\end{quote}

\subsubsection*{Response to E.2}
We have revised the manuscript to be more cautious regarding the implications and generalizability of our findings. See the revised abstract and related text edits below.

\noindent Related text edits on pages \pageref{thisdesign1}, \pageref{synsem}, \pageref{thisdesign2} and \pageref{thisdesign3}:

\begin{quote}
    \Paste{thisdesign1}
\end{quote}

\begin{quote}
    \Paste{thisdesign4}
\end{quote}

\begin{quote}
    \Paste{thisdesign2}
\end{quote}

\begin{quote}
    \Paste{thisdesign3}
\end{quote}

@@

\subsection*{Comment E.3}
\begin{quote}
``\textbf{Reviewer 2's additional point 1 that asks whether you have changed that particular analysis}.''
\end{quote}

\subsubsection*{Response to E.3}
Yes, we changed all analyses in the first revision. As we had explained in the response to reviewers on August 29, 2024 (specifically in the response letter and our responses to the comments E.2 and R3.5), the priors were changed from directional priors in the original version to symmetrical priors in the revised version. Additionally, we changed the method with which the Bayes factors were computed. Initially, we had used bridge sampling, but in the revision, we switched to the Savage-Dickey method because only using that method we were able to compare models with random slopes (this was also explained in the previous response letter, Response to R1.7). These changes did not change the results qualitatively, but led to some numerical changes which probably piqued the reviewer's interest.

\subsection*{Comment E.4}
\begin{quote}
``and \textbf{their final point about the length of the current version of the manuscript}.''
\end{quote}

\subsubsection*{Response to E.4}
We have removed repetitions in the description of the statistical analyses and have shortened the description of the self-paced reading results (and the comparison to the previous studies). This overall reduced the length of the manuscript; after the first round of revisions the main text ended on page 77, now it ends on page 73.

@@

Actually, I thought here the authors basically provided an explanation of the differences between their findings and the previous findings because of differences in the methodology, even though their last sentence said otherwise. Let me spell out it: since the current methodology (SPR, EEG) is likely more demanding on memory, it is possible that people give up on detailed processing; rather, they do something akin to good-enough processing and do not fully process; consequently, they could still get interference from semantics (since the semantic interference is really simple, it is just lexical semantics and readers don't need to fully parse for that) but they don't correctly parse embedded subjects, hence the syntactic interference is gone. \textbf{The authors in fact come to good-enough processing in Section 7.3, but then it is not linked back to differences between their findings and the other findings (which I think should be, I do think this is a possible explanation).}''
\end{quote}

\subsubsection*{Response to R1.7}
We have added the reviewer's suggested explanation that the employed methods could have especially led to good-enough processing and that this might have contributed to the different findings between studies (see page \pageref{SPR_goodenough}):

\begin{quote}
    \Paste{SPR_goodenough}
\end{quote}

\subsection*{Comment R1.8}
\begin{quote}
``p. 76: \textbf{Chromy should appear like this: ``Chromý" (note the correct diacritics above the ``y")}

@@

\section*{Reviewer \#2} 
\subsection*{Comment R2.1}
\begin{quote}
``I appreciate the authors' responses to my previous comments. However, despite the revisions, significant weaknesses remain, rendering in my opinion this study primarily a methodological contribution that adds little value to our understanding of interference effects. Additionally, the authors' strong claims often come through as overstatements that could result in serious misconceptions.
My overall assessment is as follows.

1. Interpretability of the SPR results: The results from the self-paced reading (SPR) task remain unclear. The only observable effect—the semantic interference—emerges well before the critical region of interest (specifically, at the distractor region) and persists without change beyond the critical region. As a result, any effect at the critical region is uninterpretable.

The authors acknowledge the interpretability issue with the SPR results in their response to my comment, and they now include the following statement: ``Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over regions, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. Therefore, we refrain from further discussing effects occurring in the later regions." \textbf{However, this comes after more than 10 full pages (pp. 25-37) of discussion on the reaction times (RTs) in the critical and pre-critical regions, which would lead any reader to believe these findings are meaningful. If, as the authors now realize, the effect emerges much earlier, this entire discussion is misleading: the SPR results after the distractor region are simply uninterpretable.}''
\end{quote}

\subsubsection*{Response to R2.1}
We agree with the reviewer that the presentation and discussion of the SPR results was misleading and have revised it accordingly. We have revised Figure 3 to remove the focus on the critical region (see below). At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference and that the long-lasting effect caused by that confounds the results at later effects. For comparability with previous findings, we then briefly report the later effects. We have revised the Results section in the following way, starting on page \pageref{revised_SPR}:

@@

We do not share the reviewer's interpretation that the effects in the later regions are not meaningful. We use the pre-critical reading times as a proxy to investigate the effect of encoding interference caused by the distractor. To emphasize this we have renamed the former section ``Pre-critical effects" to ``Reading times difference prior to retrieval". We have revised the beginning of the Discussion on page \pageref{prior_to_retrieval} accordingly:

\begin{quote}
    \Paste{SPR_discussion_revised}
\end{quote}

\setcounter{figure}{2}

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{blue}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom.jpg}
\end{sidewaysfigure}
\clearpage

@@

\subsubsection*{Response to R2.2}
The focus of the present study was to investigate retrieval interference due to syntactic and semantic similarity. The finding that the reading times were primarily affected by encoding interference was unexpected (but see \citeauthor{mertzen}'s (\citeyear{mertzen}) discussion). Given that the present manuscript contains two large-sample experiments and computational modeling (and its substantial length), we think that it is outside the scope of this manuscript to provide an in-depth discussion of a mechanism of encoding interference.

\subsection*{Comment R2.3}

\begin{quote}
``2. Lack of syntactic interference evidence: No evidence for syntactic interference was found in either task, contradicting two previous studies that tested similar structures (Van Dyke 2007; Mertzen et al. 2023). \textbf{In my previous review, I noted that the syntactic manipulation used was not appropriate and thus unlikely to induce interference effects. Although the authors concede that this might be true, they did not address the logical implications of this acknowledgment, as evidenced by their repeated strong assertions}, such as the one in the abstract: ``Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to syntactic cues".
If the syntactic manipulation is irrelevant, then repeatedly concluding that syntactic interference is absent is not only an overstatement but also misleading. All this study allows us to conclude is that when cues that are irrelevant to syntax are manipulated (like [+being a subject of whichever clause at whichever level of embedding], underpowered studies (Van Dyke 2007; Mertzen et al. 2023) may falsely report interference effects. While this is a valid conclusion and a sound methodological critique, it does not significantly advance our theoretical understanding of syntactic interference. Specifically, it fails to demonstrate that when relevant syntactic cues are manipulated, syntactic interference is not attested. Yet, this is the conclusion the authors consistently imply throughout their manuscript, leading to significant misconceptions. As I mentioned in my previous review, beyond the two studies the authors are building on, numerous studies in the literature show that hierarchically intervening elements do generate strong syntactic interference.
It is a fine enterprise to correct inaccurate claims from previously underpowered studies that used irrelevant manipulations; however, this should not not suggest that the findings presented here have a broader significance than they do.''
\end{quote}
\subsubsection*{Response to R2.3}
We agree with the reviewer that our present work does not rule out syntactic interference in general, but the present study using the present design did not induce considerable syntactic interference. We have revised the abstract and manuscript main body accordingly (see response to E.2).

@@

\subsection*{Comment R2.4}
\begin{quote}
``3. Misalignment of SPR and ERP results: The results from the self-paced reading (SPR) and event-related potential (ERP) methods do not align, with the only observed effect—semantic interference—pointing in opposite directions: inhibitory in SPR but facilitatory in ERP. Specifically, the authors report a reversed effect in the P600 component, showing a reduced P600 in conditions of higher semantic interference. This contradicts typical predictions, where the P600 is usually greater under high interference. The authors interpret this as a facilitatory interference effect (i.e., easier processing with high semantic interference). However, if this were the case, \textbf{we would also expect faster reaction times (RTs) in high semantic interference conditions, while the opposite is observed in the SPR data}.''
\end{quote}

\subsubsection*{Response to R2.4}
The SPR results and the ERP results in the N400 window both show inhibitory interference, i.e., more processing difficulty under high vs.\ low semantic interference, manifesting in longer reading times and a more negative N400. We explain below in our response to R2.5 that the reduced P600 for high interference might be explained by facilitatory interference because under high semantic interference whatever noun was retrieved can be easily integrated.

SPR results cannot show the same complex results pattern that the ERP results showed because they provide just one measure per region. Additionally, facilitated integration could only be observed in the critical (or spill-over) region and in line with the previous reviewer comments, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.

\subsection*{Comment R2.5}
\begin{quote}
``4. Conflicting directions in ERP components: Within the ERP results, the different components also point in opposite directions. The P600 component shows a reversed effect, with a reduced P600 in conditions of high semantic interference. However, the N400 component follows the expected pattern, with stronger activation under high semantic interference. \textbf{It remains unclear how the directionality of these two can be reconciled within the same theoretical framework.}''
\end{quote}

@@

\subsubsection*{Response to R2.5}

We have revised our discussion of the ERP results on page \pageref{inhibition_facilitation} to clarify how the seemingly conflicting results can be reconciled (see below). Interference might enhance the N400 amplitude because it makes memory retrieval more difficult and it might reduce the P600 because no matter which noun was retrieved it can be easily integrated with the verb.

@@

1. Original version: ``By contrast, the syntactic interference effect had only anecdotal evidence under the two narrow priors (Normal-(0, 0.1): BF10 = 2.5, Normal-(0, 0.5): BF10 = 1.2). There was evidence against syntactic interference under wider priors (Normal-(0, 1): BF10 = 0.66, Normal-(0, 5): BF10 = 0.14). What this means is that, in our data, only if we assume a priori that the effect size is relatively small, there is very weak evidence in favor of syntactic interference."

Revised version: ``By contrast, Bayes factors provided either no evidence for or even evidence against syntactic interference in both spatiotemporal windows (BF10 $<$ 1.2). Similarly, Bayes factors provided either no evidence for or even evidence against the interaction in both spatio-temporal windows (BF10 $<$ 1.2)."

\textbf{Did the authors change the analyses? Why what was reported as anecdotal evidence under narrow priors is now reported as no evidence or evidence against?}''
\end{quote}
\subsubsection*{Response to R2.7}
Yes, all analyses were changed in the first round of revision. Firstly, the priors were changed (directional priors in the original version and symmetrical priors in the revised version). Secondly, the method to compute Bayes factors was changed (from bridge sampling to Savage-Dickey, see also our Response to E.3). This resulted in small numerical changes of the results, e.g., BF10 = 2.5 in the original analysis decreased to BF10 = 1.2 in the revised analysis. Furthermore in the original version of the manuscript, we were inconsistent in our description of small BFs, i.e., BF10 = 1.2. Since the first revision and in line with \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence either way.

\subsection*{Comment R2.8}
\begin{quote}
``2. As a side note, on p. 58, the authors conclude that ``the parser searches for a subject that is within the same clause but uses the animacy cue without reference to the clause in which a noun appears." Does this mechanism truly seem plausible to the authors? What would this imply in practice? \textbf{Does it suggest that semantic and syntactic information are entirely encapsulated from each other?} If we were to set aside the issues I raised earlier, this would be one of the paper's central conclusions, yet the authors fail to provide a credible mechanism to support it.''
\end{quote}
\subsubsection*{Response to R2.8}

To clarify the assumed independence of syntactic and semantic processing, we have revised section 7.3 starting on page \pageref{synsem}. We have also renamed the section to `The use of syntactic and semantic information' (the former title was `Relation to other sentence processing accounts').

\begin{quote}
    \Paste{synsem}
\end{quote}

@@

\begin{quote}
``This is a revision of a paper that I previously reviewed. (I was Reviewer \#3 in the previous round.)

In the previous round of reviews, I was generally happy with the authors' findings. My queries were mostly related to the interpretation and generalizability of those findings. This is a routine issue in psycholinguistic research. A study looks at effects in one very specific (and often complicated) sentence type, and then draws conclusions about very broad notions such as `syntax' and `semantics'.

The authors have offered a thorough (daunting?) response to the reviews. I am more satisfied with some responses than others. But I do not think that this should stand in the way of publication in the JML special issue.

The authors acknowledge the concern about using relational notions such as `subject of the same clause' as a memory retrieval cue. They provide some text that clarifies that they need to implement a clause tracking mechanism. I would prefer it if the text was clearer about the unfeasibility of a `subject of the same clause' feature. But this does not impact their findings.

Another concern that I raised involves the scalability of the notion that semantic interference effects reflect the use of semantic retrieval cues such as [+animate]. At issue is how this extends to finer-grained properties that are routinely responsible for plausibility violations. In my previous review I used the example of ``the teacher drove" vs. ``the little boy drove". It is, of course, clear that teachers are more plausible drivers than young children. This is part of our world knowledge. But it is less plausible that every time the noun `teacher' is encountered the property [+can drive] is activated. The authors' response seems to be that animacy was sufficient for the materials in the current study, and that maybe other finer-grained features ``become activated only when relevant". To my mind, this point removes much of the value of the cue-based retrieval theory. \textbf{Nevertheless, this concern is about the generalizability of the authors' claims, and not about the soundness of their results.}

Finally, and in a similar vein, I raised questions about whether there might be other reasons for the selective interference effects found in this study. Maybe a different mechanism could capture the special status of the subject-of-the-same-clause, or maybe some specific property of the current experimental materials could be responsible for the lack of interference from other subjects. The authors seem skeptical of the first of these, and they seem more sympathetic to the second. \textbf{This is all reasonable enough, and their text conveys some caution. The abstract, on the other hand, is rather more confident.} I am sympathetic to the authors' preferred conclusion. But I am less confident than they are about how well justified it is based on the current findings. This has nothing to do with the quantitative wizardry that the authors display in the paper. It's all about the issue of generalizing from a single sentence configuration (that participants saw in 60\% of trials in the study).''
\end{quote}

\subsubsection*{Response to Reviewer \#3}
We have revised the manuscript (including the abstract) to show more caution regarding the generalizability of our findings (see our response to E.2).

### 2024-11-15 paper/revision2/elsarticle-template-harv.tex
Add files via upload

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiment 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_exp.jpg}
\end{sidewaysfigure}
\clearpage

\label{revised_SPR}\Copy{revised_SPR}{
It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. This pattern \textcolor{blue}{was found in the pooled data (panels A and B) as well as in the data of each experimental session (panels C and D). To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006}. Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session. As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.

@@

\paragraph{Hypothesis testing using Bayes factors}
\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}). There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5,
Normal(0, 0.05): BF$_{10}$ = 1.3,
Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). {There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).

@@

\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence} A and B). One might argue that this could be due to differences in word length. The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9). It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.

@@

{Taken together, the present SPR and ERP experiments using the 2 $\times$ 2 interference design developed by \cite{vandyke07} consistently showed effects of semantic interference and little or no effects of syntactic interference or an interaction.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy. 
The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in. A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.

In the remainder of this section, we first provide an in-depth comparison of the present results to the previous studies' results which used the same design. Then we discuss two important issues that relate to the broader interference literature. The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.

@@

{The difference in methodology led to another difference between our study and the previous studies. In their eye-tracking experiments, \cite{vandyke07} and \cite{mertzen} presented the stimuli on a single line. Whether \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment used line breaks or not is not reported. In our SPR experiments, it was necessary to introduce line breaks. These were hard-coded to occur always after the subject of the critical verb and two words after the critical verb. The line break after the subject might have increased its prominence compared to the other nouns in the sentence, i.e., the distractor. It has been proposed by \cite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability. Therefore, increased prominence of the subject could have decreased interference. Since we have found evidence for semantic, but not for syntactic interference, the increased prominence might have especially affected syntactic interference. For this point to hold, one would need to make the reasonable assumption that the line break provided additional structural information by emphasizing the syntactic structure of our sentences. However, our EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor. So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment. Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference \textcolor{blue}{in the present design}.
}\label{linebreak}

@@

The previous and present items differed in the pre-critical material. \citet{vandyke07} used one region consisting of two words (e.g., `yesterday afternoon') which were analyzed together. \cite{mertzen} used a one-word pre-critical region (e.g., `tatsächlich', indeed). In the present study, we added a pre-pre-critical region to help us differentiate between the potential sources of the pre-critical effects in the literature as discussed in the Discussion of the SPR experiments. So, the present study was the only one which had a pre-critical region that did not directly follow a clause boundary and it was also the only one to not find syntactic interference in the pre-critical region. These two observations combined suggest that the pre-critical syntactic interference effect in the previous studies might have been a spill-over effect from the syntactic manipulation. The syntactic interference manipulation in all the studies using this design was confounded with syntactic complexity of the material intervening between the subject and the (pre-)critical region. In the low syntactic interference conditions, it was a simple relative clause (who VP PP). In the high syntactic interference conditions, it was a relative clause with an embedded complement clause (who VP that NP VP). This difference in syntactic complexity could have ``spilled over'' into the pre-critical region of the previous studies causing the syntactic (interference) effect. While this seems like a plausible explanation for the previous effects, it would also apply to our pre-pre-critical region. However, Figure \ref{fig:whole_sentence} \textcolor{blue}{C} shows that the pre-pre-critical region (n-2) in our items did not show a syntactic effect.

@@

\subsection{Relation to other sentence processing accounts}

The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}. {A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference. This future work should utilize the findings from the literature on agreement attraction which has shown that the hierarchical position of the distractor can determine whether or not it affects processing \citep{franck2002subject, franck2006agreement, franck2020hierarchical, parker2018not}.}\label{future_work}

However, the semantic interference effect we found can be easily reconciled with previous work in  sentence processing. In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model.  
 Our results could also be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses. The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands. It is reasonable to assume that the participants might have -- at least occasionally -- adopted a good-enough processing mode when faced with high task demands \citep{swets2008underspecification,LogacevMultiple,LogacevVasishthQJEP2016} or due to working memory capacity limitations \citep{MalsburgVasishth2013}, or both. So, given a good-enough processing mode which leaves some syntactic relations within the sentence underspecified, it makes sense for the comprehender to use a simple heuristic for subject-verb dependency resolution. Animacy is an obvious choice for such a heuristic. Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject leads to the correct analysis with high probability (not just within the experiment context, but also in everyday language use).

@@

Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role  \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations}. However, this conclusion would be problematic. First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}. Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \citep[e.g.,][]{Avetisyan_etal_2020_agree,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and 
the importance of word order \citep[e.g.,][]{meng2000mode}.

>>>>

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{blue}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom.jpg}
\end{sidewaysfigure}
\clearpage

\label{revised_SPR}\Copy{revised_SPR}{
It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. \textcolor{blue}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006}. Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session. As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.

@@

\paragraph{Hypothesis testing using Bayes factors}
\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}). There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5, Normal(0, 0.05): BF$_{10}$ = 1.3, Normal(0, 0.1): BF$_{10}$ = 0.7). There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8). There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1). Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).

@@

\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence}). One might argue that this could be due to differences in word length. The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9). It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.

@@

{Taken together, the present SPR and ERP experiments using the 2 $\times$ 2 interference design developed by \cite{vandyke07} consistently showed effects of semantic interference and little or no effects of syntactic interference or an interaction.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy. 
The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in. \textcolor{blue}{We revisit the connection between syntax and semantics in section 7.3.} A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.

In the remainder of this section, we first provide an in-depth comparison of the present results to the previous studies' results which used the same design. Then we discuss two important issues that relate to the broader literature. The first is the role of encoding vs.\ retrieval interference, and the second is \textcolor{blue}{the use of syntactic and semantic information during sentence processing}.

@@

The difference in methodology led to another difference between our study and the previous studies. In their eye-tracking experiments, \cite{vandyke07} and \cite{mertzen} presented the stimuli on a single line. Whether \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment used line breaks or not is not reported. In our SPR experiments, it was necessary to introduce line breaks. These were hard-coded to occur always after the subject of the critical verb and two words after the critical verb. The line break after the subject might have increased its prominence compared to the other nouns in the sentence, i.e., the distractor. It has been proposed by \cite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability. Therefore, increased prominence of the subject could have decreased interference. Since we have found evidence for semantic, but not for syntactic interference, the increased prominence might have especially affected syntactic interference. For this point to hold, one would need to make the reasonable assumption that the line break provided additional structural information by emphasizing the syntactic structure of our sentences. However, our EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor. So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment. \Copy{thisdesign1}{Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference \textcolor{blue}{in the present design}.
}\label{thisdesign1}

@@

The previous and present items differed in the pre-critical material. \citet{vandyke07} used one region consisting of two words (e.g., `yesterday afternoon') which were analyzed together. \cite{mertzen} used a one-word pre-critical region (e.g., `tatsächlich', indeed). In the present study, we added a pre-pre-critical region to help us differentiate between the potential sources of the pre-critical effects in the literature as discussed in the Discussion of the SPR experiments. So, the present study was the only one which had a pre-critical region that did not directly follow a clause boundary and it was also the only one to not find syntactic interference in the pre-critical region. These two observations combined suggest that the pre-critical syntactic interference effect in the previous studies might have been a spill-over effect from the syntactic manipulation. The syntactic interference manipulation in all the studies using this design was confounded with syntactic complexity of the material intervening between the subject and the (pre-)critical region. In the low syntactic interference conditions, it was a simple relative clause (who VP PP). In the high syntactic interference conditions, it was a relative clause with an embedded complement clause (who VP that NP VP). This difference in syntactic complexity could have ``spilled over'' into the pre-critical region of the previous studies causing the syntactic (interference) effect. While this seems like a plausible explanation for the previous effects, it would also apply to our pre-pre-critical region. However, Figure \ref{fig:whole_sentence} \textcolor{blue}{C} shows that the pre-pre-critical region in our items did not show a syntactic effect.

@@

\subsection{\textcolor{blue}{The use of syntactic and semantic information}}

\label{synsem}\Copy{synsem}{\Copy{thisdesign4}{\textcolor{blue}{Our results suggest a remarkable disconnect in the use of syntactic and semantic cues during subject-verb dependency resolution in sentence configurations like the ones investigated in the present study.}} The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}. A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference. This future work should utilize the findings from the literature on agreement attraction which has shown that the hierarchical position of the distractor can determine whether or not it affects processing \citep{franck2002subject, franck2006agreement, franck2020hierarchical, parker2018not}.

\textcolor{blue}{Although the syntactic cue included hierarchical information to match only with the correct subject within the same clause, this did not block the semantic cue from matching with the embedded distractor. This finding is not consistent with syntax-first models like garden-path theory \citep{frazier1987} which assume that syntactic processing precedes all other levels of linguistic evaluation (e.g., plausibility). Our findings rather suggest parallel, fully independent matching of different retrieval cues with the candidates in memory.}  In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model \textcolor{blue}{assuming independent cues.}

The semantic interference effect we found can be easily reconciled with \textcolor{blue}{other sentence processing theories that do not assign a special status to syntactic processing}.} Our results could be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses. The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands. It is reasonable to assume that the participants might have -- at least occasionally -- adopted a good-enough processing mode when faced with high task demands \citep{swets2008underspecification,LogacevMultiple,LogacevVasishthQJEP2016} or due to working memory capacity limitations \citep{MalsburgVasishth2013}, or both. So, given a good-enough processing mode which leaves some syntactic relations within the sentence underspecified, it makes sense for the comprehender to use a simple heuristic for subject-verb dependency resolution. Animacy is an obvious choice for such a heuristic. Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject leads to the correct analysis with high probability (not just within the experiment context, but also in everyday language use).

@@

\Copy{thisdesign2}{Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role  \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations}}\label{thisdesign2}. However, this conclusion would be problematic. First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}. Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \citep[e.g.,][]{Avetisyan_etal_2020_agree,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and 
the importance of word order \citep[e.g.,][]{meng2000mode}.
